A Survey of Real-Time Strategy Game AI\\ Research and Competition in StarCraft

### 2013-07-20 survey.tex
Fixing grammatical errors and some reviewers comments

\section{Introduction}\label{sec:intro}
\IEEEPARstart{T}{he} field of real-time strategy (RTS) game AI has advanced significantly since Michael Buto's call for research in this area \cite{Buro03rts}. Specially, competitions like the ``AIIDE StarCraft AI Competition'' have motivated the exploration of many AI approaches in the context of RTS AI. We will list and classify these approaches, explain their 
strengths and weaknesses and conclude on what is left to achieve human-level 
RTS AI.

Complex dynamic environments, where it is not possible to have perfect nor complete information about the current state or about the dynamics of the environment, pose significant challenges for artificial intelligence. Road  traffic, finance, or
weather  forecasts are  examples of such large, complex, real-life dynamic
environments. RTS games can be seen as a simplification of one such real-life
environment, with simpler dynamics in a  finite and smaller
world, although still complex enough  to study some of the key interesting
problems like decision making  under
uncertainty  or  real-time adversarial planning. Finding efficient  techniques for
tackling  these problems  on  RTS  games can  thus  benefit other AI disciplines and application
domains, ans also have concrete  and direct applications in  the ever growing
industry of  video games.%, generating worldwide two  and a  half times more revenue than the Film industry in 2011.
% This statement is not fair, the reference I found it only compare all the game industry revenue (including hardware sales) vs theatrical film industry revenue (i.e. not DVD sales, ...)

@@

\begin{itemize}
\item They are {\em simultaneous move} games, where more than one player can issue actions at the same time. Additionally, these actions are {\em durative}, i.e. actions are not instantaneous, but take some amount of time time to complete.
\item RTS games are ``real-time'', which actually means is that each player has a very small amount of time to decide the next move. Compared to Chess, where players may have several minutes to decide the next action, in StarCraft, the game executes at 24 iterations per second, which means that players can act as fast as every 42ms, before the game state changes.
\item Most RTS games are partially observable: players can only see the part of the map that has been explored. This is referred to as the {\em fog-of-war}.
\item Most RTS games are non-deterministic. Some actions have a chance of success.
\item And finally, the complexity of these games, both in terms of state space size and in terms of number of actions available at each decision cycle is very large. For example, the state space of Chess is typically estimated to be around $10^{50}$, heads up no-limit texas holdem poker around $10^{80}$, and Go around $10^{170}$. In comparison, the state space of StarCraft in a typical map is estimated to be many orders of magnitude larger than any of those, as discussed in the next section.
\end{itemize}

For those reasons, standard techniques used for playing classic board games, such as game tree search, cannot be directly applied to solve RTS games without the definition of some level of abstraction, or some other simplification. Interestingly enough, humans seem to be able to deal with the complexity of RTS games, and are still vastly superior to computers in these type of games \cite{burochurchill2012aimagazine}. For those reasons, a large spectrum of techniques have been attempted to deal with this domain, as we will describe below. The remainder of this section is devoted to describe StarCraft as a research testbed, and on detailing the open challenges in RTS game AI.

@@

A typical StarCraft map is defined as a rectangular grid, where the $width \times height$ of the map is measured in the number of $32 \times 32$ square of pixels, also known as build tiles. Although the
resolution of walkable areas is in squares of $8 \times 8$ pixels, also known as walk tiles. The typical dimensions for maps range from $64 \times 64$ to $256 \times 256$ build tiles. Each player  can
control up  to 200 units (plus an unlimited number of buildings).  Moreover, each
different race  contains between 30  to 35 different types  of units
and  buildings,  each  one  with  a significant  number  of  special
abilities. All these factors  together make StarCraft a significant
challenge,   in   which   humans    are   still   much   better   than
computers. For instance, in the game ladder iCCup\footnote{http://www.iccup.com/StarCraft/} where users are ranked by their current point totals ($E$ being the lowest possible rank, and $A^+$ and $Olympic$ being the second highest and highest ranks, respectively), the best StarCraft AI bots are ranked between $D$ and $D^+$, where average  amateur players are  ranked between $C^+$ and  $B$. For comparison, StarCraft professional  players are usually ranked between
$A^-$ and $A^+$.

@@

\subsubsection{Planning}
As mentioned above, the size of the state space in RTS games is much larger than that of traditional board-games such as Chess or Go. Additionally, the number of actions that can be executed at a given instant of time is also much larger. Thus, standard adversarial planning approaches, such as game tree search are not directly applicable. As we elaborate later, planning in RTS games can be seen as having multiple levels of abstraction: at a higher level, players need long-term planning capabilities, in order to develop a strong economy in the game; at a low level, individual units need to be moved in coordination to fight battles taking into account the terrain and the opponent. Techniques that can address these large planning problems by either sampling, or hierarchical decomposition do not yet exist.

@@

\subsubsection{Domain Knowledge Exploitation}
In traditional board-games such as Chess, researchers have exploited the large amounts of existing domain knowledge to create good evaluation functions to be used by alpha-beta search algorithms, extensive opening books, or end-game tables. In the case of RTS games, it is still unclear the ways in which the significantly large amount of domain knowledge (in the forms or strategy guides, replays, etc.) can be exploited by bots. Most work in this area has focused on two main directions: on the one hand, researchers are finding ways in which to hard-code existing strategies into bots, so that bots only need to decide which strategies to deploy, instead of having to solve the complete problem of deciding which actions to execute by each individual unit at each time step. One the other hand, large datasets of replays have been created \cite{WeberCig09,synnaeve2012dataset}, from where strategies, trends or plans have been tried to learn. However, StarCraft games are quite complex, and how to automatically learn from such datasets is still an open problem.

@@

In comparison, when humans play StarCraft, they typically divide their decision making in a very different way. The StarCraft community typically talk about two tasks:
\begin{itemize}
  \item {\em Micro}: is the ability to control your units individually. A good \emph{micro} player usually keeps their units alive during more time.
  \item {\em Macro}: is the ability to produce units and to expand at the appropriate times to keep your production of units flowing. A good \emph{macro} player usually has the larger army.
\end{itemize}

@@

Figure~\ref{fig:levels-abstraction} graphically illustrates how strategy, tactics and reactive control are three points in a continuum scale where strategy corresponds to decisions making processes that affect long spans of time (several minutes in the case of StarCraft), reactive control corresponds to low-level second-by-second decisions, and tactics sit in the middle. Also, strategic decisions reason about the whole game at once, whereas tactical or reactive control decisions are localized, and affect only specific groups of units. Typically, strategic decisions constraint future tactical decisions, which in turn condition reactive control. Moreover, information gathered while performing reactive control, can cause reconsideration of the tactics being employed; which could trigger further strategical reasoning.

@@

% PLANNING:
Approaches using planning techniques have also been explored in the literature. For example Onta\~{n}\'{o}n et al. \cite{CBR_Planning} explored the use of real-time case-based planning (CBP) in the domain of Wargus (a Warcraft II clone). In their work, they used human demonstration to learn plans, that are then composed at run-time in order to form full-fledges strategies to play the game. In \cite{PlanRetrieval} they improve over  their previous CBP approach by using situation assessment for improving the quality and speed of plan retrieval. Hierarchical Task-Network (HTN) planning has also been explored with some success in the context of simpler first-person shooter games \cite{HTNPlanning}. Planning approaches offer more adaptivity of the AI strategy compared to hard-coded approaches. However, the real-time constraints of RTS games limit the planning approaches that can be applied, being HTN and case-based planning the only ones explored so far. Moreover, none of these approaches addresses any timing or scheduling issues, which are key in RTS games. On notable exception is the work of Churchill and Buro \cite{churchill2011build}, who used planning in order to construct its economic build-orders taking into account timing constraints of the different actions.

@@

% SCOUTING:
One final consideration concerning strategy is that RTS games are typically partially observable. Games like StarCraft implement the ``fog-of-war'' idea, which basically means that a player can only see the areas of the map close to her own units. Areas of the map away from the field of view of individual units are not observable. Players need to scout in order to obtain information about the opponent's strategy. The size of the state space in StarCraft prevents solutions based on POMDPs from being directly applicable, and very few of the previous approaches deal with this problem. Most research in RTS game AI assumes perfect information all the time. In the case of commercial games, most AI implementations cheat, since they have perfect information, while the human player does not. In order to make the human player believe the AI of this games does not cheat, sometimes they simulate some scouting tasks as Bob Fitch described in his AIIDE 2011 keynote for the WarCraft and StarCraft game series. A notable exception is the work of Weber et al. \cite{WeberAIIDE11}, who used a particle model with a linear trajectory update to track opponent units under fog-of-war in StarCraft. They also produced tactical goals through reactive planning and goal-driven autonomy \cite{WeberCig10,Weber10}, finding the more relevant goal(s) to spawn in unforeseen situations.

@@

% MACHINE LEARNING:
Concerning tactical decision making, many different approaches have been explored such as machine learning or game tree search. 
Hladky and Bulitko \cite{Hladky2008} benchmarked hidden semi-Markov models (HSMM) and particle filters in first person shooter games (FPS) units tracking. They showed that the accuracy of occupancy maps was improved using movement models (learned from the player behavior) in HSMM. Kabanza et al. \cite{OBRecog} improve the probabilistic hostile agent task tracker (PHATT \cite{PHATT}, a simulated HMM for plan recognition) by encoding strategies as HTN, used for plan and intent recognition to find tactical opportunities. Sharma et al. \cite{CBR-RL} combined CBR and reinforcement learning to enable reuse of tactical plan components. Cadena and Garrido \cite{CadenaG11} used fuzzy CBR (fuzzy case matching) for strategic and tactical planning. Finally, \cite{SynnaeveTactics} combined space abstraction into regions from \cite{Perkins10} and tactical-decision making by assigning scores (economical, defenses, etc.) to regions and looking for their correspondences to tactical moves (attacks) in pro-gamers replays.
% GAME TREE SEARCH:
Game tree search techniques have also been explored for tactical decision making. \cite{Chung05} applied Monte-Carlo planning to a capture-the-flag version of Open RTS. Balla and Fern \cite{UCT} applied the UCT algorithm (a Monte Carlo Tree Search algorithm) to tactical assault planning in Wargus. To make game tree search is applicable at this level, abstract game state representations are used in order to reduce the complexity. Also, abstractions, or simplifications about the set of possible actions to execute in a given game state need to be used.

@@

% POTENTIAL FIELDS:
Potential fields (or influence maps) have been found to be a useful technique for reactive decision making. Some uses of potential fields in RTS games are: avoiding obstacles (navigation), avoiding opponent fire \cite{uriarte2012kiting}, or staying at maximum shooting distance \cite{Hagelback09}. Potential fields have also been combined with A* path-finding to avoid local traps \cite{Hagelback12}. Hagelb\"{a}ck and Johansson \cite{HagelbackJ08} presented a multiagent potential fields based bot able to deal with fog-of-war in the Tankbattle game. Avery et al. \cite{Avery09} and Smith et al. \cite{SmithCIG10} co-evolved influence map trees for spatial reasoning in RTS games. Danielsiek et al. \cite{Danielsiek_2008} used influence maps to achieve intelligent squad movement to flank the opponent in a RTS game. Despite their success, a drawback for potential field-based techniques is the large number of parameters that has to be tuned in order to achieve the desired behavior. Approaches for automatically learning such parameters have been explored, for example, using reinforcement learning \cite{Liu_2008}, or self-organizing-maps (SOM) \cite{teamCompositionRTS}. We would like to note that potential fields are a reactive control technique, and as such, they do not perform any form of lookahead. As a consequence, these techniques are prone to make units stuck in local maxima.

@@

This section reviews the results of the recent international competitions on AI for StarCraft. These competitions, typically co-located with scientific conferences, have been possible thanks to the existence of the the 
Brood War Application Programming Interface 
(BWAPI)\footnote{\url{http://code.google.com/p/bwapi/}}, which 
enables replacing the human player interface with C++ code.
%A consequence of this indirect way of integrating new bots is that every machine can only run one custom bot, requiring two computers for a 2-player game.
The following subsections summarize the results of all the StarCraft AI competitions held at the AIIDE (Artificial Intelligence for Interactive Digital Entertainment) and CIG (Computational Intelligence in Games) conferences during the past years. Additionally we analyze the statistics from the StarCraft Bot Ladder, where the best bots play against each other continuously over time.

@@

Tournament 4 was the complete game of {\em StarCraft: Brood War} with fog-of-war enforced. The tournament
was run with a random pairing double-elimination format with each match being best of 5 games.
Competitors could play as any of the three races, with the only limitations in gameplay
being those which were considered ``cheating'' in the StarCraft community. A map pool of
5 well-known professional maps was announced to competitors in advance, with a random map being chosen for each game.

Results are shown in Table \ref{tab:aiide2010}. The team that won was Overmind\footnote{\url{http://overmind.cs.berkeley.edu}}, from University of California, Berkeley. Using the Zerg race, their
strategy was to defend early aggression with zergling units while amassing mutalisk units
which they used to contain and eventually defeat their opponents. The mutalisk is a very fast and agile
flying unit which is able to attack while moving with no drawback, which makes them quite a powerful
unit when controlled by a computer. Overmind used a potential-field
based micro-management system to guide their mutalisks, which led them to victory. Krasi0 came in 2nd place
with a standard defensive Terran opening strategy which transitioned into mech play in the late game.

@@

In 2011 the competition was hosted by the University of Alberta, with organization by Michael Buro and
David Churchill\footnote{\url{https://skatgame.net/mburo/sc2011/}}. Due to a lack of entrants in tournament categories 1-3 in the 2010 competition, it was
decided that only the full game category would be played in the 2011 competition. Another important
change in the 2011 competition was the introduction of automated tournament-managing software running
StarCraft games simultaneously on 20 computers, allowing a total of 1170 games to be played in far less
time than the 108 games of the 2010 competition. This increase in games played also allowed the tournament
to switch to a round-robin format, eliminating the ``luck'' factor of the pairings inherent in bracket
style tournaments. The bot which achieved the highest win percentage over the course of the competition would
be determined the winner. Also, the competition became open-source, in an effort not
only to prevent possible cheating, but to promote healthy competition in future tournaments by giving
newcomers and easier entry point by basing their design off of previous bots.

@@

UAlbertaBot used an early zealot-rush strategy to take advantage of the power of early game Protoss units. 
It would send out the first zealots which were made and immediately attack the enemy base, using a unit
counting heuristic to determine whether or retreat or keep pushing. Of note is that UAlbertaBot used an
online planning algorithm to construct all of its economic build-orders\cite{churchill2011build}, as no hard-coded build orders
were used.

AIUR also chose  Protoss, with a strategy which  was in between Skynet
and UAlbertaBot in terms of  attack timings.  At that time, AIUR chose
one  mood among  five (leading  to slightly  different  strategies and
tactics) at the  beginning of a game and kept it  until the end. These
five moods were:
\begin{itemize}
\item  Rush:  where  the  bot  tries  early  attacks,  and  have  good
  probabilities to send the two or three first Zealots (basic contact attack ground unit) to harass the
  opponent.
\item Aggressive: where  we have less chance to  perform harasses with
  the first Zealots, and the first attack is usually a bit delayed
  with regard to the Rush mood.
\item Macro: where the AI do not try any early attacks and focus a bit
  more on its economy before attacking.
\item Defense: where the AI  ``turtles'' and wait to have a consequent
  army before running an attack.
\item  Fast expand:  where the  first building  constructed it  a base
  expansion, for a very economical-oriented game.
\end{itemize}
Notice that  build orders are not  fully hard-coded since  they can be
altered by AIUR's Spend Manager.

@@

The 2012 competition was also hosted by the University of Alberta, with the major difference from the 2011
competition being the addition of persistent storage. Bots could now write information to disk during a
match, and then read the information during other matches, allowing them to adjust strategies based
on previous results. 6 of the 10 entrants used this feature to aid in strategy selection, including the
top 4 finishers. More improvements to the tournament environment also meant that a total of 4240 games
could now be played in the same time period. Results are shown in Table \ref{tab:aiide2012}.

@@

In this match it was shown that the true weakness of state of the art StarCraft AI systems was that
humans are very adept at recognizing scripted behaviors and exploiting them to the fullest. A human
player in Skynet's position in the first game would have realized he was being taken advantage of and
adapted his strategy accordingly, however the inability to put the local context (Bakuryu kiting his
units around his base) into the larger context of the game (that this would delay Skynet until
reinforcements arrived) and then the lack of strategy change to fix the situation led to an easy
victory for the human. These problems remain as some of the the main challenges in RTS AI today: to both recognize
the strategy and intent of an opponent's actions, and how to effectively adapt your own strategy to 
overcome them.

@@

The overall results are displayed in table~\ref{tab:cig2012},
and the win rate evolution over time in figure~\ref{fig:cig2012-results}.
These are quite consistent with the results of the AIIDE 2012 competition,
so that we can conclude that the best bots are not very dependent on 
knowing the maps beforehand. However, the bot vs. bot win rates as
displayed in figure~\ref{fig:cig2012-mapresults} show some interesting
trends. On the maps with more possible start points, some bots do 
better than others, namely SCAIL, Adjutant, Nova, and UAlbertaBot,
the latter probably due to its very efficient scouting routine.
Some bots however suffer from the increased uncertainty about
the enemies position, namely Xelnaga and BroodwarBotQ.

@@

The StarCraft Bot Ladder is a website\footnote{\url{http://bots-stats.krasi0.com}} where bot versus bot matches are automatized, and are running all the time. This ladder is a great resource for creating data sets (all the game replays are available) and statistics. For bot ranking, the ladder uses an Elo rating system suitable for calculating the \emph{relative skill level} of a bot in two-player games. In the Elo system each player has a numerical rating that gets incremented or decremented some points after each game. The amount of points depends on the difference in the ratings of the players. A player will gain more points by beating a higher-rated player than by beating a lower-rated player. These kind of rating system is widely used in games like chess. This bot ladder compiles different versions of bots from the main worldwide competitions (like AIIDE, CIG or SSCAI\footnote{\url{http://sscaitournament.com}}), even some independent or ``under construction'' bots. Therefore, it is a very good resource to test the performance of new bots against the current state of the art in StarCraft bots before participating in the official competitions.

>>>>

\section{Introduction}\label{sec:intro}
\IEEEPARstart{T}{he} field of real-time strategy (RTS) game AI has advanced significantly since Michael Buro's call for research in this area \cite{Buro03rts}. Specially, competitions like the ``AIIDE StarCraft AI Competition'' have motivated the exploration of many AI approaches in the context of RTS AI. We will list and classify these approaches, explain their 
strengths and weaknesses and conclude on what is left to achieve human-level 
RTS AI.

Complex dynamic environments, where it is not possible to have neither perfect nor complete information about the current state or about the dynamics of the environment, pose significant challenges for artificial intelligence. Road  traffic, finance, or
weather  forecasts are  examples of such large, complex, real-life dynamic
environments. RTS games can be seen as a simplification of one such real-life
environment, with simpler dynamics in a  finite and smaller
world, although still complex enough  to study some of the key interesting
problems like decision making  under
uncertainty  or  real-time adversarial planning. Finding efficient  techniques for
tackling  these problems  on  RTS  games can  thus  benefit other AI disciplines and application
domains, and also have concrete  and direct applications in  the ever growing
industry of  video games.%, generating worldwide two  and a  half times more revenue than the Film industry in 2011.
% This statement is not fair, the reference I found it only compare all the game industry revenue (including hardware sales) vs theatrical film industry revenue (i.e. not DVD sales, ...)

@@

\begin{itemize}
\item They are {\em simultaneous move} games, where more than one player can issue actions at the same time. Additionally, these actions are {\em durative}, i.e. actions are not instantaneous, but take some amount of time to complete.
\item RTS games are ``real-time'', which actually means is that each player has a very small amount of time to decide the next move. Compared to Chess, where players may have several minutes to decide the next action, in StarCraft, the game executes at 24 iterations per second, which means that players can act as fast as every 42ms, before the game state changes.
\item Most RTS games are partially observable: players can only see the part of the map that has been explored. This is referred to as the {\em fog-of-war}.
\item Most RTS games are non-deterministic. Some actions have a chance of success.
\item And finally, the complexity of these games, both in terms of state space size and in terms of number of actions available at each decision cycle is very large. For example, the state space of Chess is typically estimated to be around $10^{50}$, heads up no-limit Texas holdem poker around $10^{80}$, and Go around $10^{170}$. In comparison, the state space of StarCraft in a typical map is estimated to be many orders of magnitude larger than any of those, as discussed in the next section.
\end{itemize}

For those reasons, standard techniques used for playing classic board games, such as game tree search, cannot be directly applied to solve RTS games without the definition of some level of abstraction, or some other simplification. Interestingly enough, humans seem to be able to deal with the complexity of RTS games, and are still vastly superior to computers in these types of games \cite{burochurchill2012aimagazine}. For those reasons, a large spectrum of techniques have been attempted to deal with this domain, as we will describe below. The remainder of this section is devoted to describe StarCraft as a research testbed, and on detailing the open challenges in RTS game AI.

@@

A typical StarCraft map is defined as a rectangular grid, where the $width \times height$ of the map is measured in the number of $32 \times 32$ square of pixels, also known as build tiles. Although the
resolution of walkable areas is in squares of $8 \times 8$ pixels, also known as walk tiles. The typical dimensions for maps range from $64 \times 64$ to $256 \times 256$ build tiles. Each player  can
control up  to 200 units (plus an unlimited number of buildings).  Moreover, each
different race  contains between 30  to 35 different types  of units
and  buildings,  most of them with  a significant  number  of  special
abilities. All these factors  together make StarCraft a significant
challenge,   in   which   humans    are   still   much   better   than
computers. For instance, in the game ladder iCCup\footnote{http://www.iccup.com/StarCraft/} where users are ranked by their current point totals ($E$ being the lowest possible rank, and $A^+$ and $Olympic$ being the second highest and highest ranks, respectively), the best StarCraft AI bots are ranked between $D$ and $D^+$, where average  amateur players are  ranked between $C^+$ and  $B$. For comparison, StarCraft professional  players are usually ranked between
$A^-$ and $A^+$.

@@

\subsubsection{Planning}
As mentioned above, the size of the state space in RTS games is much larger than that of traditional board games such as Chess or Go. Additionally, the number of actions that can be executed at a given instant of time is also much larger. Thus, standard adversarial planning approaches, such as game tree search are not directly applicable. As we elaborate later, planning in RTS games can be seen as having multiple levels of abstraction: at a higher level, players need long-term planning capabilities, in order to develop a strong economy in the game; at a low level, individual units need to be moved in coordination to fight battles taking into account the terrain and the opponent. Techniques that can address these large planning problems by either sampling, or hierarchical decomposition do not yet exist.

@@

\subsubsection{Domain Knowledge Exploitation}
In traditional board games such as Chess, researchers have exploited the large amounts of existing domain knowledge to create good evaluation functions to be used by alpha-beta search algorithms, extensive opening books, or end-game tables. In the case of RTS games, it is still unclear the ways in which the significantly large amount of domain knowledge (in the forms or strategy guides, replays, etc.) can be exploited by bots. Most work in this area has focused on two main directions: on the one hand, researchers are finding ways in which to hard-code existing strategies into bots, so that bots only need to decide which strategies to deploy, instead of having to solve the complete problem of deciding which actions to execute by each individual unit at each time step. One the other hand, large datasets of replays have been created \cite{WeberCig09,synnaeve2012dataset}, from where strategies, trends or plans have been tried to learn. However, StarCraft games are quite complex, and how to automatically learn from such datasets is still an open problem.

@@

In comparison, when humans play StarCraft, they typically divide their decision making in a very different way. The StarCraft community typically talks about two tasks:
\begin{itemize}
  \item {\em Micro}: is the ability to control your units individually. A good \emph{micro} player usually keeps their units alive during more time.
  \item {\em Macro}: is the ability to produce units and to expand at the appropriate times to keep your production of units flowing. A good \emph{macro} player usually has the larger army.
\end{itemize}

@@

Figure~\ref{fig:levels-abstraction} graphically illustrates how strategy, tactics and reactive control are three points in a continuum scale where strategy corresponds to decisions making processes that affect long spans of time (several minutes in the case of StarCraft), reactive control corresponds to low-level second-by-second decisions, and tactics sit in the middle. Also, strategic decisions reason about the whole game at once, whereas tactical or reactive control decisions are localized, and affect only specific groups of units. Typically, strategic decisions constraint future tactical decisions, which in turn condition reactive control. Moreover, information gathered while performing reactive control, can cause reconsideration of the tactics being employed; which could trigger further strategic reasoning.

@@

% PLANNING:
Approaches using planning techniques have also been explored in the literature. For example Onta\~{n}\'{o}n et al. \cite{CBR_Planning} explored the use of real-time case-based planning (CBP) in the domain of Wargus (a Warcraft II clone). In their work, they used human demonstration to learn plans, which are then composed at run-time in order to form full-fledges strategies to play the game. In \cite{PlanRetrieval} they improve over  their previous CBP approach by using situation assessment for improving the quality and speed of plan retrieval. Hierarchical Task-Network (HTN) planning has also been explored with some success in the context of simpler first-person shooter games \cite{HTNPlanning}. Planning approaches offer more adaptivity of the AI strategy compared to hard-coded approaches. However, the real-time constraints of RTS games limit the planning approaches that can be applied, being HTN and case-based planning the only ones explored so far. Moreover, none of these approaches addresses any timing or scheduling issues, which are key in RTS games. On notable exception is the work of Churchill and Buro \cite{churchill2011build}, who used planning in order to construct its economic build-orders taking into account timing constraints of the different actions.

@@

% SCOUTING:
One final consideration concerning strategy is that RTS games are typically partially observable. Games like StarCraft implement the ``fog-of-war'' idea, which basically means that a player can only see the areas of the map close to her own units. Areas of the map away from the field of view of individual units are not observable. Players need to scout in order to obtain information about the opponent's strategy. The size of the state space in StarCraft prevents solutions based on POMDPs from being directly applicable, and very few of the previous approaches deal with this problem. Most research in RTS game AI assumes perfect information all the time, this mean that they only consider the information that they can see to make decisions. In the case of commercial games, most AI implementations cheat, since they have perfect information, while the human player does not. In order to make the human player believe the AI of these games does not cheat, sometimes they simulate some scouting tasks as Bob Fitch described in his AIIDE 2011 keynote for the WarCraft and StarCraft game series. A notable exception is the work of Weber et al. \cite{WeberAIIDE11}, who used a particle model with a linear trajectory update to track opponent units under fog-of-war in StarCraft. They also produced tactical goals through reactive planning and goal-driven autonomy \cite{WeberCig10,Weber10}, finding the more relevant goal(s) to spawn in unforeseen situations.

@@

% MACHINE LEARNING:
Concerning tactical decision making, many different approaches have been explored such as machine learning or game tree search. 
Hladky and Bulitko \cite{Hladky2008} benchmarked hidden semi-Markov models (HSMM) and particle filters in first person shooter games (FPS) units tracking. They showed that the accuracy of occupancy maps was improved using movement models (learned from the player behavior) in HSMM. Kabanza et al. \cite{OBRecog} improve the probabilistic hostile agent task tracker (PHATT \cite{PHATT}, a simulated HMM for plan recognition) by encoding strategies as HTN, used for plan and intent recognition to find tactical opportunities. Sharma et al. \cite{CBR-RL} combined CBR and reinforcement learning to enable reuse of tactical plan components. Cadena and Garrido \cite{CadenaG11} used fuzzy CBR (fuzzy case matching) for strategic and tactical planning. Finally, \cite{SynnaeveTactics} combined space abstraction into regions from \cite{Perkins10} and tactical-decision making by assigning scores (economical, defenses, etc.) to regions and looking for their correspondences to tactical moves (attacks) in pro-gamers replays.
% GAME TREE SEARCH:
Game tree search techniques have also been explored for tactical decision making. \cite{Chung05} applied Monte-Carlo planning to a capture-the-flag version of Open RTS. Balla and Fern \cite{UCT} applied the UCT algorithm (a Monte Carlo Tree Search algorithm) to tactical assault planning in Wargus. To make game tree search applicable at this level, abstract game state representations are used in order to reduce the complexity. Also, abstractions, or simplifications about the set of possible actions to execute in a given game state need to be used.

@@

% POTENTIAL FIELDS:
Potential fields and influence maps have been found to be a useful technique for reactive decision making. Some uses of potential fields in RTS games are: avoiding obstacles (navigation), avoiding opponent fire \cite{uriarte2012kiting}, or staying at maximum shooting distance \cite{Hagelback09}. Potential fields have also been combined with A* path-finding to avoid local traps \cite{Hagelback12}. Hagelb\"{a}ck and Johansson \cite{HagelbackJ08} presented a multiagent potential fields based bot able to deal with fog-of-war in the Tankbattle game. Avery et al. \cite{Avery09} and Smith et al. \cite{SmithCIG10} co-evolved influence map trees for spatial reasoning in RTS games. Danielsiek et al. \cite{Danielsiek_2008} used influence maps to achieve intelligent squad movement to flank the opponent in a RTS game. Despite their success, a drawback for potential field-based techniques is the large number of parameters that has to be tuned in order to achieve the desired behavior. Approaches for automatically learning such parameters have been explored, for example, using reinforcement learning \cite{Liu_2008}, or self-organizing-maps (SOM) \cite{teamCompositionRTS}. We would like to note that potential fields are a reactive control technique, and as such, they do not perform any form of lookahead. As a consequence, these techniques are prone to make units stuck in local maxima.

@@

This section reviews the results of the recent international competitions on AI for StarCraft. These competitions, typically co-located with scientific conferences, have been possible thanks to the existence of the Brood War Application Programming Interface 
(BWAPI)\footnote{\url{http://code.google.com/p/bwapi/}}, which 
enables replacing the human player interface with C++ code.
%A consequence of this indirect way of integrating new bots is that every machine can only run one custom bot, requiring two computers for a 2-player game.
The following subsections summarize the results of all the StarCraft AI competitions held at the AIIDE (Artificial Intelligence for Interactive Digital Entertainment) and CIG (Computational Intelligence in Games) conferences during the past years. Additionally we analyze the statistics from the StarCraft Bot Ladder, where the best bots play against each other continuously over time.

@@

Tournament 4 was the complete game of {\em StarCraft: Brood War} with fog-of-war enforced. The tournament
was run with a random pairing double-elimination format with each match being best of 5 games.
Competitors could play as any of the three races, with the only limitations in gameplay
being those that were considered ``cheating'' in the StarCraft community. A map pool of
5 well-known professional maps were announced to competitors in advance, with a random map being chosen for each game.

Results are shown in Table \ref{tab:aiide2010}. The team that won was Overmind\footnote{\url{http://overmind.cs.berkeley.edu}}, from University of California, Berkeley. Using the Zerg race, their
strategy was to defend early aggression with zergling units while amassing mutalisk units,
which they used to contain and eventually defeat their opponents. The mutalisk is a very fast and agile
flying unit which is able to attack while moving with no drawback, which makes them quite a powerful
unit when controlled by a computer. Overmind used a potential-field
based micro-management system to guide their mutalisks, which led them to victory. Krasi0 came in 2nd place
with a standard defensive Terran opening strategy that transitioned into mech play in the late game.

@@

In 2011 the University of Alberta hosted the competition, with organization by Michael Buro and
David Churchill\footnote{\url{https://skatgame.net/mburo/sc2011/}}. Due to a lack of entrants in tournament categories 1-3 in the 2010 competition, it was
decided that only the full game category would be played in the 2011 competition. Another important
change in the 2011 competition was the introduction of automated tournament-managing software running
StarCraft games simultaneously on 20 computers, allowing a total of 1170 games to be played in far less
time than the 108 games of the 2010 competition. This increase in games played also allowed the tournament
to switch to a round-robin format, eliminating the ``luck'' factor of the pairings inherent in bracket
style tournaments. The bot that achieved the highest win percentage over the course of the competition would
be determined the winner. Also, the competition became open-source, in an effort not
only to prevent possible cheating, but to promote healthy competition in future tournaments by giving
newcomers and easier entry point by basing their design off of previous bots.

@@

UAlbertaBot used an early zealot-rush strategy to take advantage of the power of early game Protoss units. 
It would send out the first zealots that were made and immediately attack the enemy base, using a unit
counting heuristic to determine whether or retreat or keep pushing. Of note is that UAlbertaBot used an
online planning algorithm to construct all of its economic build-orders\cite{churchill2011build}, as no hard-coded build orders
were used.

AIUR also chose  Protoss, with a strategy that was in between Skynet
and UAlbertaBot in terms of  attack timings.  At that time, AIUR chose
one  mood among  five (leading  to slightly  different  strategies and
tactics) at the  beginning of a game and kept it  until the end. These
five moods were:
\begin{itemize}
\item  Rush:  where  the  bot  tries  early  attacks,  and  have  good
  probabilities to send the two or three first Zealots (basic contact attack ground unit) to harass the
  opponent.
\item Aggressive: where  we have less chance to  perform harasses with
  the first Zealots, and the first attack is usually a bit delayed
  with regard to the Rush mood.
\item Macro: where the AI do not try any early attacks and focus a bit
  more on its economy before attacking.
\item Defense: where the AI  ``turtles'' and wait to have a consequent
  army before running an attack.
\item  Fast expand:  where the  first building  constructed it  a base
  expansion, for a very economical-oriented game.
\end{itemize}
Notice that  build orders are not  fully hard-coded since  they can be
altered by AIUR's Spend Manager.

@@

The University of Alberta also hosted the 2012 competition, with the major difference from the 2011
competition being the addition of persistent storage. Bots could now write information to disk during a
match, and then read the information during other matches, allowing them to adjust strategies based
on previous results. 6 of the 10 entrants used this feature to aid in strategy selection, including the
top 4 finishers. More improvements to the tournament environment also meant that a total of 4240 games
could now be played in the same time period. Results are shown in Table \ref{tab:aiide2012}.

@@

In this match it was shown that the true weakness of state of the art StarCraft AI systems was that
humans are very adept at recognizing scripted behaviors and exploiting them to the fullest. A human
player in Skynet's position in the first game would have realized he was being taken advantage of and
adapted his strategy accordingly, however the inability to put the local context (Bakuryu kiting his
units around his base) into the larger context of the game (that this would delay Skynet until
reinforcements arrived) and then the lack of strategy change to fix the situation led to an easy
victory for the human. These problems remain as some of the main challenges in RTS AI today: to both recognize
the strategy and intent of an opponent's actions, and how to effectively adapt your own strategy to 
overcome them.

@@

The overall results are displayed in table~\ref{tab:cig2012},
and the win rate evolution over time in figure~\ref{fig:cig2012-results}.
These are quite consistent with the results of the AIIDE 2012 competition,
so that we can conclude that the best bots are not very dependent on 
knowing the maps beforehand. However, the bot vs. bot win rates as
displayed in figure~\ref{fig:cig2012-mapresults} show some interesting
trends. On the maps with more possible start points, some bots do 
better than others, namely SCAIL, Adjutant, Nova, and UAlbertaBot,
the latter probably due to its very efficient scouting routine.
Some bots however suffer from the increased uncertainty about
the enemies' position, namely Xelnaga and BroodwarBotQ.

@@

The StarCraft Bot Ladder is a website\footnote{\url{http://bots-stats.krasi0.com}} where bot versus bot matches are automatized, and are running all the time. This ladder is a great resource for creating data sets (all the game replays are available) and statistics. For bot ranking, the ladder uses an Elo rating system suitable for calculating the \emph{relative skill level} of a bot in two-player games. In the Elo system each player has a numerical rating that gets incremented or decremented some points after each game. The amount of points depends on the difference in the ratings of the players. A player will gain more points by beating a higher-rated player than by beating a lower-rated player. This kind of rating system is widely used in games like chess. This bot ladder compiles different versions of bots from the main worldwide competitions (like AIIDE, CIG or SSCAI\footnote{\url{http://sscaitournament.com}}), even some independent or ``under construction'' bots. Therefore, it is a very good resource to test the performance of new bots against the current state of the art in StarCraft bots before participating in the official competitions.
